---
title: "ML_State_Tree"
author: "Xin Feng"
date: "2020/2/25"
output:
  pdf_document: default
  html_document:
    df_print: paged
geometry: margin=0.75in
fontsize: 12pt
fig_crop: no
---


```{r setup, include=FALSE}
library(gridExtra)
library(tidyverse)
library(mosaic)
library(broom)
library(modelr)
library(car)
library(knitr)
library(GGally)
library(dummies)
library(MASS)
library(foreign)
library(ISLR)
library(rsample)
library(rcfss)
library(yardstick)
library(tree)
library(e1071)
library(class) # for kNN
library(arm) # for a few base features
library(partykit)
library(reldist) # gini coefficient
options(width=70, digits=4, scipen=8)
knitr::opts_chunk$set(size='small') # Set the default R output size a bit smaller
```


##Loading Data
```{r}
df1 <- read.csv("MLData_Clean.csv")
```

##Variables: 
#HTOTVAL: household income
#A_SEX: 1 = Male, 2 = Female
#A_AGE
#PRCITSHP: 1 = Native, born in US; 2 = Native, born in PR or US outlying area; 3 = Native, born abroad of US parent(s); 4 = Foreign born, US cit by naturalization; 5 = Foreign born, not a US citizen
#PRDTRACE: 1 = White only, 2 = Black only, 3 = AI only, 4 = Asian only, 5 = HP only, 6 = Mixture (Clean)
#PEHSPNON: Are you Spanish, Hispanic, or Latino?  1 = Yes 2 = No
#PENATVTY: Place of Birth 1 = U.S., 2 = Outside of U.S. (Clean)
#PEFNTVTY: Father's Place of Birth 1 = U.S., 2 = Outside of U.S. (Clean)
#PEMNTVTY: Mother's Place of Birth 1 = U.S., 2 = Outside of U.S. (Clean)
#PEPAR1TYP (DAD): -1 = Dead, 1 = Biological, 2 = Step, 3 = Adopted
#PEPAR2TYP (MOM): -1 = Dead, 1 = Biological, 2 = Step, 3 = Adopted
#P_STAT Values: 1 = Civilian 15+, 2 = Armed Forces
#A_MJOCC: Major Occupation Code 0-11
#A_MJIND: Major Industry Code 0-14

#Variables and Factor
```{r}
# merged data
df1s <- select(df1, "HTOTVAL","A_SEX","PRCITSHP","PRDTRACE","PENATVTY","PEFNTVTY","PEMNTVTY","PEPAR1TYP","PEPAR2TYP","P_STAT", "A_MJOCC", "A_MJIND", "GESTFIPS", "state", "state_full")
# converting interger to factors
df1s[sapply(df1s, is.integer)] <- lapply(df1s[sapply(df1s, is.integer)], as.factor)
df1s[,1] <- as.integer(df1s[,1])/1000
head(df1s)
```

##PART I: 
##Filter the dataset by state 1-56
##Within each state, further divide the data into training/testing set
##Then, run a ctree on training set and then a ctree on the testing set
##We use the tree result to generate MSE and gini coefficient


##Separate the states, recreate a list consistes of 51 states. (5 states do not have data)
```{r}
state_length <- 56
state_list <- vector(mode = "list", length = state_length)
for (i in 1:state_length) {
    state_list[[i]] <- df1s %>% filter(GESTFIPS == i)
}

#5 states we don't have any record, remove them
state_list[[3]] <- NULL
state_list[[6]] <- NULL
state_list[[12]] <- NULL
state_list[[40]] <- NULL
state_list[[48]] <- NULL
```


##Training and Testing Set List
```{r}
train_length <- 51
test_length <- 51
train_list <- vector(mode = "list", length = 51)
test_list <- vector(mode = "list", length = 51)

set.seed(100)
for (i in 1:train_length) {
  num = seq(from=1, to=nrow(state_list[[i]]), by=1) 
  train_list[[i]]=sample(1:nrow(state_list[[i]]),
                           nrow(state_list[[i]])*0.75)
}

for (j in 1:test_length) {
    k = 1
    test_list[[j]] = rep("NA", times = nrow(state_list[[j]])*0.25)
    num = seq(from=1, to=nrow(state_list[[j]]), by=1) 
    for (m in 1:length(num)) {
       if(is.element(num[m], train_list[[j]]) == FALSE){
         test_list[[j]][k] = num[m]
         k = k + 1
       }
    }
    test_list[[j]] = as.integer(test_list[[j]])
    test_list[[j]] = sample(test_list[[j]])
}

```


##Tree List
```{r}
tree_list <- vector(mode = "list", length = 51)

for (i in 1:length(tree_list)) {
 train <- train_list[[i]]
 citree <- ctree(HTOTVAL ~ .,
                data=state_list[[i]][train, ],
                control = ctree_control(alpha=0.05,
                                        testtype="Bonferroni"))
 tree_list[[i]] <- citree
}
```


##Graph
```{r, fig.width=30, fig.height=20}
# Plot
#for (i in 1:2){
plot(tree_list[[1]])
#}
```

##MES and Opportunity based Gini coefficient of 51 states
```{r}
# mse
mse <- seq(from=1, to=51, by=1)
for (i in 1:51) {
  test <- test_list[[i]]
  mse[i] <- mean((state_list[[i]][test, ]$HTOTVAL - 
  predict(tree_list[[i]],newdata = state_list[[i]][test, ]))^2)
}

# opportunity based gini coefficient
gini_opp <- seq(from=1, to=51, by=1)
for (i in 1:51) {
  gini_opp[i] <- gini(predict(tree_list[[i]],newdata = state_list[[i]]))
}

```